

# Lets look at twitter data!

## Preamble: Load libs, read data

Load libraries. `lubridate` and `scales` deal with time data. `grid` allows plotting side by side `ggplot2` objects. 
```{r libs}
library(ggplot2)
library(reshape2)
library(dplyr)
library(magrittr)
library(lubridate)
library(grid)
library(scales)
```

Read in data downloaded from [analytics.twitter.com](analytical.twitter.com

```{r readDat}
setwd('~/Dropbox/Documents/statsforbios/twitterData')
d <- read.csv('tweet_activity_metrics0121.csv', stringsAsFactors = FALSE)

dim(d)
names(d)
```

Set `ggplot2` options.

```{r ggplotDefaults}
theme_set(theme_minimal(base_size = 20))

```



## Some basic time stuff

Number of tweets has remained relatively constant over the period.

```{r time}
d$time %<>% as.POSIXct
ggplot(d, aes(x = time)) +
  geom_histogram(bin = 7*24*60*60) +
  ylab('No. of tweets (weekly)') +
  xlab('Date')
```

Most of my tweets are between 10 and 5 GMT. Quite a few in the evening as well.

```{r timeOnly}
d$timeOnly <- hour(d$time) + minute(d$time)/60

ggplot(d, aes(x = timeOnly)) + 
  geom_histogram() +
  ylab('No. of tweets') +
  xlab('Time (hr)')
```


## Impressions exploratory analysis

Another reminder of first rule of stats. Plot the data.

```{r impr}
imprPlot <- ggplot(d, aes(x = time, y = impressions)) +
  geom_point(alpha = 0.3) +
  scale_y_sqrt() +
  scale_x_datetime("", labels = date_format("%b"))

rtPlot <- ggplot(d, aes(x = time, y = retweets)) + 
  geom_point(alpha = 0.3) +
  scale_x_datetime("", labels = date_format("%b"))

pushViewport(viewport(layout = grid.layout(1, 2)))
print(imprPlot, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(rtPlot, vp = viewport(layout.pos.row = 1, layout.pos.col = 2))

```

Turns out this dataset has all my tweets, but only records impressions from October 2014 (when I started using twitter analytics. 
So, any analysis on impressions must be done only on the later data. 
Retweets are fine.
I'm just going to remove all old tweets. 
Seems cleaner.

```{r oldOnly}
d %<>% filter(., time > 1409150040)
```

## Replies

Add a logical column of whether a tweet is a reply.

```{r plotReplies}
d$reply <- factor(substring(d$Tweet.text, 1, 1) == '@')

ggplot(d, aes(y = impressions, x = reply)) +
  geom_violin() +
  scale_y_log10()
```

Unsurprisingly, there are much fewer impressions for replies. 
So I'm going to remove them from further analysis as well.

```{r rmReplies}
d %<>% filter(reply == FALSE)
```

## Time regressions

Time of day is marginally significant, but not much there.
```{r impreVtime}
lm(d$impressions ~ d$timeOnly) %>% summary

ggplot(d, aes(x = timeOnly, y = impressions)) +
  geom_point() +
  scale_y_sqrt()

```

However I think it's slightly better to do from 6am rather than midnight.

```{r rotateTime, warning = FALSE, message = FALSE}
d$frmMorn <- d$timeOnly - 6
d$frmMorn[d$frmMorn < 0] <- d$frmMorn[d$frmMorn < 0] + 24 

ggplot(d, aes(x = frmMorn)) + 
  geom_histogram()

ggplot(d, aes(x = frmMorn, y = impressions)) +
  geom_point() +
  scale_y_sqrt() +
  stat_smooth(method = 'lm', se = FALSE, size = 1.7, linetype = '22') + 
  stat_smooth(method = 'glm', family = 'quasipoisson', colour = 'red4', se = FALSE, size = 1.7, linetype = '21')


lm(d$impressions ~ d$frmMorn) %>% summary
glm(d$impressions ~ d$frmMorn, family = 'quasipoisson') %>% summary
```

Linear model is not significant. 
GLM is not significant. 
I don't think there's anything here. 


### Try removing retweets

In a sense, when looking at factors such as time of day, we want to look at the number of impressions *irrespective* of how good that tweet was. 
So I'll remove the retweeted tweets and redo analysis.

```{r rmRT}
rmRT <- d %>% filter(retweets == 0)


ggplot(rmRT, aes(x = frmMorn, y = impressions)) +
  geom_point() +
  scale_y_sqrt() +
  stat_smooth(method = 'lm', se = FALSE, size = 1.7, linetype = '22') + 
  stat_smooth(method = 'glm', family = 'quasipoisson', colour = 'red4', se = FALSE, size = 1.7, linetype = '21')


lm(rmRT$impressions ~ rmRT$frmMorn) %>% summary
glm(rmRT$impressions ~ rmRT$frmMorn, family = 'quasipoisson') %>% summary

```


## Pictures

Turns out the data doesn't indicate whether a tweet contains an image.
There's an app [here](http://stackoverflow.com/questions/6500721/find-where-a-t-co-link-goes-to) that turns `t.co` domains back into original urls.
Could probably use this and then `grep` for `.png`, `.jpg` etc. etc.
But that's a pain. So for now I'm just going to use `embedded.media.clicks > 0` as a proxy for 'contains embedded media.'
This obviously biases the number of impressions upwards.

```{r embedMedia}

ggplot(d, aes(x = embedded.media.clicks > 0, y = impressions)) +
  geom_violin() +
  scale_y_log10()

glm(d$impressions ~ (d$embedded.media.clicks > 0), family='quasipoisson') %>% summary
```

So while significant, the effect size isn't massive.
Including a picture increases impressions by 50% before accounting for the bias mentioned above.

## Hashtags

Do hashtags affect the number of people that see my tweets? 
I'll look at `#dataviz` and `#rstats` as these are the only hashtags I use a lot.

```{r hashtags}
ggplot(d, aes(x = grepl('#rstats', d$Tweet.text), y = impressions)) + 
  geom_violin() + 
  scale_y_log10() +
  xlab('#rstats')

glm(d$impressions ~ (grepl('#rstats', d$Tweet.text)), family='quasipoisson') %>% summary

ggplot(d, aes(x = grepl('#dataviz', d$Tweet.text), y = impressions)) + 
  geom_violin() + 
  scale_y_log10() +
  xlab('#dataviz')

glm(d$impressions ~ (grepl('#dataviz', d$Tweet.text)), family='quasipoisson') %>% summary
```

So `#rstats` does not increase the number of impressions while `#dataviz` does by about 50%.
However there's a few tweets with very large numbers of impressions that are probably driving most of the `#dataviz` effect.


