# Power of different linear models

If you want to test for a change over a variable we can consider some different ways to do it.

We can collect the data so that it is spread out along the x axis or clumped at either end.

We can analyse the x axis as a continuous variable or a discrete variable (binning the x if it is spread out.)
These four options are displayed below with boxplots implying the data is analysed as discrete x values.
Note that only two sets of data are simulated, one clumped and one spread out.

```{r libs, echo = FALSE}
library(ggplot2)
library(magrittr)
library(reshape2)
```

```{r examplePlots, fig.show = 'hold', out.width = 350, warning = FALSE, echo = FALSE}
# Spread out data
x1 <- runif(30)
y1 <- x1 * 2 + rnorm(30)

# Clumped data
x2 <- rep(c(0,1), 15)
y2 <- x2 * 2 + rnorm(30)
data <- data.frame(x1 = x1, x2 = x2, y1 = y1, y2 = y2)

ggplot(data, aes(x = x1, y = y1)) +
  geom_point() + 
  stat_smooth(method = 'lm', se = FALSE) +
  ggtitle('Spread out, continuous x')


ggplot(data, aes(x = x1, y = y1 )) +
  geom_boxplot(aes(group = (x1 > 0.5)*1, y = y1, colour = factor((x1 > 0.5)*1)),  weight = 2) +
  geom_point(aes(colour = factor((x1 > 0.5)*1))) +
  theme(legend.position="none") +
  ggtitle('Spread out, binned and discrete x')

ggplot(data, aes(x = x2, y = y2)) +
  geom_point() + 
  stat_smooth(method = 'lm', se = FALSE) +
  ggtitle('Clumped, continuous x')

ggplot(data, aes(x = factor(x2), y = y2, colour = factor(x2))) +
  geom_boxplot() + 
  geom_point() +
  theme(legend.position="none") +
  ggtitle('Clumped, discrete x')



```

So to examine the power of these approaches here's a function that simulates some data (from a linear model with normal error) and then calculates and extracts p-values (sorry) for the four cases shown above.


```{r lmFunc}
calcLM <- function(){

x1 <- runif(30)
y1 <- x1 * 2 + rnorm(30)
x2 <- rep(c(0,1), 15)
y2 <- x2 * 2 + rnorm(30)

coef <- c( summary(lm(y1 ~ x1))$coef[8],
           summary(lm(y2 ~ x2))$coef[8],
           summary(lm(y1 ~ x1 > 0.5))$coef[8],
           summary(lm(y2 ~ as.factor(x2)))$coef[8]
          )
}
```

Then let's run the simulation 1000 times.

```{r sim, cache = TRUE, warning = FALSE, message = FALSE}
p <- t(replicate(1000, calcLM())) %>% data.frame

colnames(p) <- c('continuousSpread', 'continuousClumped', 'discreteSpread', 'discreteClumped')

pLong <- melt(p, variable.name = 'model', value.name = 'p')
```

And plot the results.
```{r simPlot, out.width = 500}
ggplot(pLong, aes(x = model, y = p)) +
  scale_y_log10() +
  geom_violin() +
  ggtitle('p values from different models.')
```

So, using data from the edges of our range of x values gives us more power (lower p-values). 
Also, it's interesting to note that doing a discrete ANOVA with the x values as a factor is identical to treating this as a continuous linear model.
This will not be true if you have more than two groups though. 
Furthermore, if you actually want to use this as a linear model, you will have to do an extra step to scale the coefficients if you do an ANOVA rather than a linear model.

So... that was kinda fun. And another chance to get to know ggplot2 better. Some code is suppressed here but you can see the full knitr document [here](lmVSanova.Rmd).






